---
title: "PredictingBreakouts"
output: html_document
date: "2023-08-22"
---

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(reshape2) 
```

We've noticed that the model does particularly bad at predicting breakouts - players who's score jumped by more than 5-10 points. Lets see if we can pull any from those in the past and see where we get to

-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

Data Loading and Cleaning from WRScorePredictionModel.Rmd, redone.

```{r}
data = read.csv("../Data/WR_Advanced_Stats - PivotData_Clean.csv")
head(data)
```
Create table that has player column for a single year, with the deltas of the previous year
Yards columns loaded in as character, so change them 
```{r}
data = data %>%
  mutate_at(c('X2017_Yards', 'X2018_Yards', 'X2019_Yards', 'X2020_Yards', 'X2021_Yards', 'X2022_Yards'), readr::parse_number)
```

```{r}
wrscore_21_22 = data %>%
  select(Player, 
         X2021_Team, X2021_Yards, X2021_YardsRun, X2021_Targets, X2021_YardsPerRoute, X2021_OpenScore, X2021_CatchScore, X2021_YACScore, X2021_Overall,
         X2022_Team, X2022_Yards, X2022_YardsRun, X2022_Targets, X2022_YardsPerRoute, X2022_OpenScore, X2022_CatchScore, X2022_YACScore, X2022_Overall
         ) %>%
  mutate(
    team_change = 1 * (X2021_Team == X2022_Team), 
    yards_delta = X2022_Yards - X2021_Yards, 
    yardsrun_delta = X2022_YardsRun - X2021_YardsRun,
    targets_delta = X2022_Targets - X2021_Targets,
    yardsperroute_delta = X2022_YardsPerRoute - X2021_YardsPerRoute,
    openscore_delta = X2022_OpenScore - X2021_OpenScore,
    catchscore_delta = X2022_CatchScore - X2021_CatchScore,
    yacscore_delta = X2022_YACScore - X2021_YACScore,
    overall_delta = X2022_Overall - X2021_Overall) %>%
  na.exclude() %>%
  mutate(Player = paste(Player, "21-22")) %>%
  rename(Y0_Yards = X2021_Yards,
         Y1_Yards = X2022_Yards,
         Y0_YardsRun = X2021_YardsRun,
         Y1_YardsRun = X2022_YardsRun,
         Y0_Targets = X2021_Targets,
         Y1_Targets = X2022_Targets,
         Y0_YardsPerRoute = X2021_YardsPerRoute,
         Y1_YardsPerRoute = X2022_YardsPerRoute,
         Y0_CatchScore = X2021_CatchScore,
         Y1_CatchScore = X2022_CatchScore,
         Y0_OpenScore = X2021_OpenScore,
         Y1_OpenScore = X2022_OpenScore,
         Y0_YACScore = X2021_YACScore,
         Y1_YACScore = X2022_YACScore,
         Y0_Overall = X2021_Overall,
         Y1_Overall = X2022_Overall) %>%
  select(-X2021_Team, -X2022_Team)
   

wrscore_20_21 = data %>%
  select(Player, 
         X2021_Team, X2021_Yards, X2021_YardsRun, X2021_Targets, X2021_YardsPerRoute, X2021_OpenScore, X2021_CatchScore, X2021_YACScore, X2021_Overall,
         X2020_Team, X2020_Yards, X2020_YardsRun, X2020_Targets, X2020_YardsPerRoute, X2020_OpenScore, X2020_CatchScore, X2020_YACScore, X2020_Overall
         ) %>%
  mutate(
    team_change = 1 * (X2021_Team == X2020_Team), 
    yards_delta = X2021_Yards - X2020_Yards, 
    yardsrun_delta = X2021_YardsRun - X2020_YardsRun,
    targets_delta = X2021_Targets - X2020_Targets,
    yardsperroute_delta = X2021_YardsPerRoute - X2020_YardsPerRoute,
    openscore_delta = X2021_OpenScore - X2020_OpenScore,
    catchscore_delta = X2021_CatchScore - X2020_CatchScore,
    yacscore_delta = X2021_YACScore - X2020_YACScore,
    overall_delta = X2021_Overall - X2020_Overall) %>%
  na.exclude() %>%
  mutate(Player = paste(Player, "20-21")) %>%
  rename(Y0_Yards = X2020_Yards,
         Y1_Yards = X2021_Yards,
         Y0_YardsRun = X2020_YardsRun,
         Y1_YardsRun = X2021_YardsRun,
         Y0_Targets = X2020_Targets,
         Y1_Targets = X2021_Targets,
         Y0_YardsPerRoute = X2020_YardsPerRoute,
         Y1_YardsPerRoute = X2021_YardsPerRoute,
         Y0_CatchScore = X2020_CatchScore,
         Y1_CatchScore = X2021_CatchScore,
         Y0_OpenScore = X2020_OpenScore,
         Y1_OpenScore = X2021_OpenScore,
         Y0_YACScore = X2020_YACScore,
         Y1_YACScore = X2021_YACScore,
         Y0_Overall = X2020_Overall,
         Y1_Overall = X2021_Overall) %>%
  select(-X2021_Team, -X2020_Team)

wrscore_19_20 = data %>%
  select(Player, 
         X2019_Team, X2019_Yards, X2019_YardsRun, X2019_Targets, X2019_YardsPerRoute, X2019_OpenScore, X2019_CatchScore, X2019_YACScore, X2019_Overall,
         X2020_Team, X2020_Yards, X2020_YardsRun, X2020_Targets, X2020_YardsPerRoute, X2020_OpenScore, X2020_CatchScore, X2020_YACScore, X2020_Overall
         ) %>%
  mutate(
    team_change = 1 * (X2019_Team == X2020_Team), 
    yards_delta = X2020_Yards - X2019_Yards, 
    yardsrun_delta = X2020_YardsRun - X2019_YardsRun,
    targets_delta = X2020_Targets - X2019_Targets,
    yardsperroute_delta = X2020_YardsPerRoute - X2019_YardsPerRoute,
    openscore_delta = X2020_OpenScore - X2019_OpenScore,
    catchscore_delta = X2020_CatchScore - X2019_CatchScore,
    yacscore_delta = X2020_YACScore - X2019_YACScore,
    overall_delta = X2020_Overall - X2019_Overall) %>%
  na.exclude() %>%
  mutate(Player = paste(Player, "19-20")) %>%
  rename(Y0_Yards = X2019_Yards,
         Y1_Yards = X2020_Yards,
         Y0_YardsRun = X2019_YardsRun,
         Y1_YardsRun = X2020_YardsRun,
         Y0_Targets = X2019_Targets,
         Y1_Targets = X2020_Targets,
         Y0_YardsPerRoute = X2019_YardsPerRoute,
         Y1_YardsPerRoute = X2020_YardsPerRoute,
         Y0_CatchScore = X2019_CatchScore,
         Y1_CatchScore = X2020_CatchScore,
         Y0_OpenScore = X2019_OpenScore,
         Y1_OpenScore = X2020_OpenScore,
         Y0_YACScore = X2019_YACScore,
         Y1_YACScore = X2020_YACScore,
         Y0_Overall = X2019_Overall,
         Y1_Overall = X2020_Overall) %>%
  select(-X2019_Team, -X2020_Team)

wrscore_18_19 = data %>%
  select(Player, 
         X2019_Team, X2019_Yards, X2019_YardsRun, X2019_Targets, X2019_YardsPerRoute, X2019_OpenScore, X2019_CatchScore, X2019_YACScore, X2019_Overall,
         X2018_Team, X2018_Yards, X2018_YardsRun, X2018_Targets, X2018_YardsPerRoute, X2018_OpenScore, X2018_CatchScore, X2018_YACScore, X2018_Overall
         ) %>%
  mutate(
    team_change = 1 * (X2019_Team == X2018_Team), 
    yards_delta = X2019_Yards - X2018_Yards, 
    yardsrun_delta = X2019_YardsRun - X2018_YardsRun,
    targets_delta = X2019_Targets - X2018_Targets,
    yardsperroute_delta = X2019_YardsPerRoute - X2018_YardsPerRoute,
    openscore_delta = X2019_OpenScore - X2018_OpenScore,
    catchscore_delta = X2019_CatchScore - X2018_CatchScore,
    yacscore_delta = X2019_YACScore - X2018_YACScore,
    overall_delta = X2019_Overall - X2018_Overall) %>%
  na.exclude() %>%
  mutate(Player = paste(Player, "18-19")) %>%
  rename(Y0_Yards = X2018_Yards,
         Y1_Yards = X2019_Yards,
         Y0_YardsRun = X2018_YardsRun,
         Y1_YardsRun = X2019_YardsRun,
         Y0_Targets = X2018_Targets,
         Y1_Targets = X2019_Targets,
         Y0_YardsPerRoute = X2018_YardsPerRoute,
         Y1_YardsPerRoute = X2019_YardsPerRoute,
         Y0_CatchScore = X2018_CatchScore,
         Y1_CatchScore = X2019_CatchScore,
         Y0_OpenScore = X2018_OpenScore,
         Y1_OpenScore = X2019_OpenScore,
         Y0_YACScore = X2018_YACScore,
         Y1_YACScore = X2019_YACScore,
         Y0_Overall = X2018_Overall,
         Y1_Overall = X2019_Overall) %>%
  select(-X2019_Team, -X2018_Team)

wrscore_17_18 = data %>%
  select(Player, 
         X2017_Team, X2017_Yards, X2017_YardsRun, X2017_Targets, X2017_YardsPerRoute, X2017_OpenScore, X2017_CatchScore, X2017_YACScore, X2017_Overall,
         X2018_Team, X2018_Yards, X2018_YardsRun, X2018_Targets, X2018_YardsPerRoute, X2018_OpenScore, X2018_CatchScore, X2018_YACScore, X2018_Overall
         ) %>%
  mutate(
    team_change = 1 * (X2017_Team == X2018_Team), 
    yards_delta = X2018_Yards - X2017_Yards, 
    yardsrun_delta = X2018_YardsRun - X2017_YardsRun,
    targets_delta = X2018_Targets - X2017_Targets,
    yardsperroute_delta = X2018_YardsPerRoute - X2017_YardsPerRoute,
    openscore_delta = X2018_OpenScore - X2017_OpenScore,
    catchscore_delta = X2018_CatchScore - X2017_CatchScore,
    yacscore_delta = X2018_YACScore - X2017_YACScore,
    overall_delta = X2018_Overall - X2017_Overall) %>%
  na.exclude() %>%
  mutate(Player = paste(Player, "17-18")) %>%
  rename(Y0_Yards = X2017_Yards,
         Y1_Yards = X2018_Yards,
         Y0_YardsRun = X2017_YardsRun,
         Y1_YardsRun = X2018_YardsRun,
         Y0_Targets = X2017_Targets,
         Y1_Targets = X2018_Targets,
         Y0_YardsPerRoute = X2017_YardsPerRoute,
         Y1_YardsPerRoute = X2018_YardsPerRoute,
         Y0_CatchScore = X2017_CatchScore,
         Y1_CatchScore = X2018_CatchScore,
         Y0_OpenScore = X2017_OpenScore,
         Y1_OpenScore = X2018_OpenScore,
         Y0_YACScore = X2017_YACScore,
         Y1_YACScore = X2018_YACScore,
         Y0_Overall = X2017_Overall,
         Y1_Overall = X2018_Overall) %>%
  select(-X2017_Team, -X2018_Team)



fulldata_deltas = rbind(wrscore_21_22, wrscore_20_21, wrscore_19_20, wrscore_18_19, wrscore_17_18)
```


```{r}
fulldata_deltas$overall_delta_abs = abs(fulldata_deltas$overall_delta)
big_changes = 
  fulldata_deltas %>%
  arrange(-overall_delta_abs) %>%
  filter(overall_delta_abs >= 10)

set.seed(1491)

sample <- sample(c(TRUE, FALSE), nrow(x_var), replace=TRUE, prob=c(0.8,0.2))
train <- fulldata_deltas[sample, ]
test <- fulldata_deltas[!sample, ]
```


```{r}
wrscore.lm <- lm(formula = Y1_Overall ~ Y0_Yards + Y0_YardsRun + Y0_Targets + Y0_YardsPerRoute + Y0_CatchScore + Y0_OpenScore + Y0_YACScore + Y0_Overall + team_change,
                 data = train)

print(summary(wrscore.lm))


wrscore.lm <- lm(formula = Y1_Overall ~ Y0_Yards + Y0_CatchScore + Y0_Overall,
                 data = train)

print(summary(wrscore.lm))

```

Try K-Means to see if there's clusters 

```{r}
library(cluster)    # clustering algorithms
library(factoextra) # clustering algorithms & visualization

fulldata_deltas_scaled <- fulldata_deltas
fulldata_deltas_scaled[,c(2:27)] <- scale(fulldata_deltas_scaled[,c(2:27)])
fulldata_deltas_scaled
```

```{r}
fulldata_deltas_scaled <- na.omit(fulldata_deltas_scaled)
k2 <- kmeans(fulldata_deltas_scaled[,c(2:27)], centers = 5, nstart = 25)
str(k2)
```

Find optimal clusters
```{r}
set.seed(502)

fviz_nbclust(fulldata_deltas_scaled[,c(2:27)], kmeans, method = "wss")

```

seems like 4 is optimal 

```{r}
k4 <- kmeans(fulldata_deltas_scaled[,c(2:27)], centers = 4, nstart = 25)
str(k4)
```


```{r}
table(k4$cluster)

fulldata_deltas$k4_cluster = k4$cluster

fulldata_deltas %>%
  group_by(k4_cluster) %>%
  summarise_if(is.numeric, mean)

fviz_cluster(k4, data = fulldata_deltas_scaled[,c(2:27)])
```

Cluster 3 shows to be the breakouts! Low Y0 numbers, with a noticable jump in yards, yardsperroute, open score, and catchscore, with also the only groups to increase in yards per route and YAC score, and overall WR score
- 

```{r}
fulldata_deltas %>%
  filter(k4_cluster == 3) %>%
  arrange(-Y0_Yards)
```


Create clusters, see with just Y0 data, see if theres any patterns


```{r}

set.seed(502)
fviz_nbclust(fulldata_deltas_scaled[,c(2:9, 18)], kmeans, method = "wss", k.max = 15)

fviz_nbclust(fulldata_deltas_scaled[,c(2:9, 18)], kmeans, method = "silhouette")


gap_stat <- clusGap(fulldata_deltas_scaled[,c(2:9, 18)], FUN = kmeans, nstart = 25,
                    K.max = 10, B = 50)
fviz_gap_stat(gap_stat)
```

While elbow is definetly at 2 clusters, at 5 clusters we see a significant slope change in all 3 - so we'll use that for more groups.

```{r}
k0 <- kmeans(fulldata_deltas_scaled[,c(2:9, 18)], centers = 5, nstart = 25)
str(k0)
```


```{r}
fulldata_deltas$k0_cluster = k0$cluster
fulldata_deltas %>%
  group_by(k0_cluster) %>%
  summarise_if(is.numeric, mean)
```
- only one group had a significant average overall_delta increase
- increase in yards and yards per route, but on average decrease in targets
--- had the lowest yards per route year 0, and did not have an increase in production
- all stayed on the same team

```{r}
fviz_cluster(k0, data = fulldata_deltas_scaled[,c(2:9, 18)])
```


Specifically, those with a team change seem to have the most breakout potential
```{r}
fulldata_deltas %>%
  filter(k0_cluster == 1) %>%
  ggplot(aes(overall_delta)) +
  geom_histogram(binwidth = 5)

fulldata_deltas %>%
  filter(k0_cluster == 1) %>%
  ggplot(aes(overall_delta)) +
  geom_histogram(binwidth = 5, color = 'blue', fill = 'lightblue') + 
  theme_minimal() + 
  labs(title = "Overall Score Delta Player Breakouts via KMeans Clustering",
         subtitle = "Players in Cluster 1 who stayed on the same team, >50% increased by at least 10, others steady",
         x = "Overall Score Delta", 
         y = "Number of Players") 

ggsave("../Images/BreakoutCluster_Histogram.png", bg = 'white')


fulldata_deltas %>%
  filter(k0_cluster == 1) %>%
  ggplot(aes(Y0_Overall, Y1_Overall)) +
  geom_point() +
  geom_text_repel(aes(label=ifelse(overall_delta>15,as.character(Player),''))) +
  geom_abline(slope=1, intercept = 0, color = 'black') +
  geom_text(aes(40, 90, label = "Overall Score Increased",), color = 'lightgreen') +
  geom_text(aes(60, 30, label = "Overall Score Decreased"), color = 'red') +
    labs(title = "WR Score Breakout Players Y0 vs Y1 Overall",
         subtitle = "Super Breakout Players (Increase of 15+) labeled",
         x = "Year 0 WR Score", 
         y = "Year 1 WR Score") +
  theme_minimal() +
    xlim(20, 80) +
    ylim(20, 90)

ggsave("../Images/BreakoutCluster_Y0vY1Overalls.png", bg = 'white')
```

Now with 24 breakouts identified, lets see if we can find the 2022 WRs that were closest to them, to predict the breakouts 

```{r}
teamchange_2022 = read.csv("../Data/2022TeamChange.csv", header = FALSE, col.names = c('Player', 'TeamChange'))
teamchange_2022
```

```{r}
data_2022 = 
data %>%
  select(c(1,48:55)) %>%
  na.omit() %>%
  mutate(team_change = teamchange_2022$TeamChange)

colnames(data_2022) = c("Player", colnames(fulldata_deltas_scaled[,c(2:9, 18)]))
data_2022_scaled = data_2022
data_2022_scaled[,c(2:10)] = scale(data_2022[,c(2:10)])



#install.packages("flexclust")
library("flexclust")
set.seed(502)
cl1 = kcca(fulldata_deltas_scaled[,c(2:9, 18)], k=5, kccaFamily("kmeans"))
cl1

pred_train <- predict(cl1)
pred_test <- predict(cl1, newdata = data_2022_scaled[,c(2:10)])

image(cl1)
points(fulldata_deltas_scaled[,c(2:9, 18)], col=pred_train, pch=19, cex=0.3)
points(data_2022_scaled[,c(2:10)], col=pred_test, pch=22, bg="orange")
```

```{r}
fulldata_deltas$cl_cluster = pred_train

fulldata_deltas %>%
  group_by(k0_cluster, cl_cluster) %>%
  count()

fulldata_deltas %>%
  group_by(cl_cluster) %>%
  summarise_if(is.numeric, mean)

fulldata_deltas %>%
  filter(cl_cluster == 4) %>%
  ggplot(aes(overall_delta)) +
  geom_histogram(binwidth = 5)

fulldata_deltas %>%
  select(Player, Y0_Overall, Y1_Overall, overall_delta, k0_cluster, cl_cluster) %>%
  arrange(-overall_delta)
```

Cluster 1 in original Kmeans is aligned with cluster 4 here, so we can use this to predict
```{r}
data_2022$cl_cluster = pred_test

data_2022 %>%
  filter(cl_cluster == 4) %>%
  arrange(-Y0_Yards)
```

24 Names to have in increase in Score, lets see what aligns with the top 10 increases (regardless of cluster), and see where we end up. We are looking for 

- young players with ambigious roles in past years 
- an above average amount of yards (~600-700)
- low yards run (~400)
- extremely low catchscore (20-30)

```{r}
fulldata_deltas %>%
  arrange(-overall_delta) %>%
  head(20) %>%
  summarise_if(is.numeric, mean, na.rm = TRUE)
```
Top WR Breakouts
Rebounds:
- Gabe Davis
- Marquez Valdes-Scantling
Ascensions:
- Alec Pierce 
- Romeo Doubs	

