---
title: "CatchScorePredictionModels"
output: html_document
date: "2023-08-15"
---



```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(reshape2) 

```


```{r}
data = read.csv("WR_Advanced_Stats - PivotData_Clean.csv")
head(data)
```

Create table that has player column for a single year, with the deltas of the previous year

Yards columns loaded in as character, so change them 
```{r}
data = data %>%
  mutate_at(c('X2017_Yards', 'X2018_Yards', 'X2019_Yards', 'X2020_Yards', 'X2021_Yards', 'X2022_Yards'), readr::parse_number)
```

```{r}
wrscore_21_22 = data %>%
  select(Player, 
         X2021_Team, X2021_Yards, X2021_YardsRun, X2021_Targets, X2021_YardsPerRoute, X2021_OpenScore, X2021_CatchScore, X2021_YACScore, X2021_Overall,
         X2022_Team, X2022_Yards, X2022_YardsRun, X2022_Targets, X2022_YardsPerRoute, X2022_OpenScore, X2022_CatchScore, X2022_YACScore, X2022_Overall
         ) %>%
  mutate(
    team_change = 1 * (X2021_Team == X2022_Team), 
    yards_delta = X2022_Yards - X2021_Yards, 
    yardsrun_delta = X2022_YardsRun - X2021_YardsRun,
    targets_delta = X2022_Targets - X2021_Targets,
    yardsperroute_delta = X2022_YardsPerRoute - X2021_YardsPerRoute,
    openscore_delta = X2022_OpenScore - X2021_OpenScore,
    catchscore_delta = X2022_CatchScore - X2021_CatchScore,
    yacscore_delta = X2022_YACScore - X2021_YACScore,
    overall_delta = X2022_Overall - X2021_Overall) %>%
  na.exclude() %>%
  mutate(Player = paste(Player, "21-22")) %>%
  rename(Y0_Yards = X2021_Yards,
         Y1_Yards = X2022_Yards,
         Y0_YardsRun = X2021_YardsRun,
         Y1_YardsRun = X2022_YardsRun,
         Y0_Targets = X2021_Targets,
         Y1_Targets = X2022_Targets,
         Y0_YardsPerRoute = X2021_YardsPerRoute,
         Y1_YardsPerRoute = X2022_YardsPerRoute,
         Y0_CatchScore = X2021_CatchScore,
         Y1_CatchScore = X2022_CatchScore,
         Y0_OpenScore = X2021_OpenScore,
         Y1_OpenScore = X2022_OpenScore,
         Y0_YACScore = X2021_YACScore,
         Y1_YACScore = X2022_YACScore,
         Y0_Overall = X2021_Overall,
         Y1_Overall = X2022_Overall) %>%
  select(-X2021_Team, -X2022_Team)
   

wrscore_20_21 = data %>%
  select(Player, 
         X2021_Team, X2021_Yards, X2021_YardsRun, X2021_Targets, X2021_YardsPerRoute, X2021_OpenScore, X2021_CatchScore, X2021_YACScore, X2021_Overall,
         X2020_Team, X2020_Yards, X2020_YardsRun, X2020_Targets, X2020_YardsPerRoute, X2020_OpenScore, X2020_CatchScore, X2020_YACScore, X2020_Overall
         ) %>%
  mutate(
    team_change = 1 * (X2021_Team == X2020_Team), 
    yards_delta = X2021_Yards - X2020_Yards, 
    yardsrun_delta = X2021_YardsRun - X2020_YardsRun,
    targets_delta = X2021_Targets - X2020_Targets,
    yardsperroute_delta = X2021_YardsPerRoute - X2020_YardsPerRoute,
    openscore_delta = X2021_OpenScore - X2020_OpenScore,
    catchscore_delta = X2021_CatchScore - X2020_CatchScore,
    yacscore_delta = X2021_YACScore - X2020_YACScore,
    overall_delta = X2021_Overall - X2020_Overall) %>%
  na.exclude() %>%
  mutate(Player = paste(Player, "20-21")) %>%
  rename(Y0_Yards = X2020_Yards,
         Y1_Yards = X2021_Yards,
         Y0_YardsRun = X2020_YardsRun,
         Y1_YardsRun = X2021_YardsRun,
         Y0_Targets = X2020_Targets,
         Y1_Targets = X2021_Targets,
         Y0_YardsPerRoute = X2020_YardsPerRoute,
         Y1_YardsPerRoute = X2021_YardsPerRoute,
         Y0_CatchScore = X2020_CatchScore,
         Y1_CatchScore = X2021_CatchScore,
         Y0_OpenScore = X2020_OpenScore,
         Y1_OpenScore = X2021_OpenScore,
         Y0_YACScore = X2020_YACScore,
         Y1_YACScore = X2021_YACScore,
         Y0_Overall = X2020_Overall,
         Y1_Overall = X2021_Overall) %>%
  select(-X2021_Team, -X2020_Team)

wrscore_19_20 = data %>%
  select(Player, 
         X2019_Team, X2019_Yards, X2019_YardsRun, X2019_Targets, X2019_YardsPerRoute, X2019_OpenScore, X2019_CatchScore, X2019_YACScore, X2019_Overall,
         X2020_Team, X2020_Yards, X2020_YardsRun, X2020_Targets, X2020_YardsPerRoute, X2020_OpenScore, X2020_CatchScore, X2020_YACScore, X2020_Overall
         ) %>%
  mutate(
    team_change = 1 * (X2019_Team == X2020_Team), 
    yards_delta = X2020_Yards - X2019_Yards, 
    yardsrun_delta = X2020_YardsRun - X2019_YardsRun,
    targets_delta = X2020_Targets - X2019_Targets,
    yardsperroute_delta = X2020_YardsPerRoute - X2019_YardsPerRoute,
    openscore_delta = X2020_OpenScore - X2019_OpenScore,
    catchscore_delta = X2020_CatchScore - X2019_CatchScore,
    yacscore_delta = X2020_YACScore - X2019_YACScore,
    overall_delta = X2020_Overall - X2019_Overall) %>%
  na.exclude() %>%
  mutate(Player = paste(Player, "19-20")) %>%
  rename(Y0_Yards = X2019_Yards,
         Y1_Yards = X2020_Yards,
         Y0_YardsRun = X2019_YardsRun,
         Y1_YardsRun = X2020_YardsRun,
         Y0_Targets = X2019_Targets,
         Y1_Targets = X2020_Targets,
         Y0_YardsPerRoute = X2019_YardsPerRoute,
         Y1_YardsPerRoute = X2020_YardsPerRoute,
         Y0_CatchScore = X2019_CatchScore,
         Y1_CatchScore = X2020_CatchScore,
         Y0_OpenScore = X2019_OpenScore,
         Y1_OpenScore = X2020_OpenScore,
         Y0_YACScore = X2019_YACScore,
         Y1_YACScore = X2020_YACScore,
         Y0_Overall = X2019_Overall,
         Y1_Overall = X2020_Overall) %>%
  select(-X2019_Team, -X2020_Team)

wrscore_18_19 = data %>%
  select(Player, 
         X2019_Team, X2019_Yards, X2019_YardsRun, X2019_Targets, X2019_YardsPerRoute, X2019_OpenScore, X2019_CatchScore, X2019_YACScore, X2019_Overall,
         X2018_Team, X2018_Yards, X2018_YardsRun, X2018_Targets, X2018_YardsPerRoute, X2018_OpenScore, X2018_CatchScore, X2018_YACScore, X2018_Overall
         ) %>%
  mutate(
    team_change = 1 * (X2019_Team == X2018_Team), 
    yards_delta = X2019_Yards - X2018_Yards, 
    yardsrun_delta = X2019_YardsRun - X2018_YardsRun,
    targets_delta = X2019_Targets - X2018_Targets,
    yardsperroute_delta = X2019_YardsPerRoute - X2018_YardsPerRoute,
    openscore_delta = X2019_OpenScore - X2018_OpenScore,
    catchscore_delta = X2019_CatchScore - X2018_CatchScore,
    yacscore_delta = X2019_YACScore - X2018_YACScore,
    overall_delta = X2019_Overall - X2018_Overall) %>%
  na.exclude() %>%
  mutate(Player = paste(Player, "18-19")) %>%
  rename(Y0_Yards = X2018_Yards,
         Y1_Yards = X2019_Yards,
         Y0_YardsRun = X2018_YardsRun,
         Y1_YardsRun = X2019_YardsRun,
         Y0_Targets = X2018_Targets,
         Y1_Targets = X2019_Targets,
         Y0_YardsPerRoute = X2018_YardsPerRoute,
         Y1_YardsPerRoute = X2019_YardsPerRoute,
         Y0_CatchScore = X2018_CatchScore,
         Y1_CatchScore = X2019_CatchScore,
         Y0_OpenScore = X2018_OpenScore,
         Y1_OpenScore = X2019_OpenScore,
         Y0_YACScore = X2018_YACScore,
         Y1_YACScore = X2019_YACScore,
         Y0_Overall = X2018_Overall,
         Y1_Overall = X2019_Overall) %>%
  select(-X2019_Team, -X2018_Team)

wrscore_17_18 = data %>%
  select(Player, 
         X2017_Team, X2017_Yards, X2017_YardsRun, X2017_Targets, X2017_YardsPerRoute, X2017_OpenScore, X2017_CatchScore, X2017_YACScore, X2017_Overall,
         X2018_Team, X2018_Yards, X2018_YardsRun, X2018_Targets, X2018_YardsPerRoute, X2018_OpenScore, X2018_CatchScore, X2018_YACScore, X2018_Overall
         ) %>%
  mutate(
    team_change = 1 * (X2017_Team == X2018_Team), 
    yards_delta = X2018_Yards - X2017_Yards, 
    yardsrun_delta = X2018_YardsRun - X2017_YardsRun,
    targets_delta = X2018_Targets - X2017_Targets,
    yardsperroute_delta = X2018_YardsPerRoute - X2017_YardsPerRoute,
    openscore_delta = X2018_OpenScore - X2017_OpenScore,
    catchscore_delta = X2018_CatchScore - X2017_CatchScore,
    yacscore_delta = X2018_YACScore - X2017_YACScore,
    overall_delta = X2018_Overall - X2017_Overall) %>%
  na.exclude() %>%
  mutate(Player = paste(Player, "17-18")) %>%
  rename(Y0_Yards = X2017_Yards,
         Y1_Yards = X2018_Yards,
         Y0_YardsRun = X2017_YardsRun,
         Y1_YardsRun = X2018_YardsRun,
         Y0_Targets = X2017_Targets,
         Y1_Targets = X2018_Targets,
         Y0_YardsPerRoute = X2017_YardsPerRoute,
         Y1_YardsPerRoute = X2018_YardsPerRoute,
         Y0_CatchScore = X2017_CatchScore,
         Y1_CatchScore = X2018_CatchScore,
         Y0_OpenScore = X2017_OpenScore,
         Y1_OpenScore = X2018_OpenScore,
         Y0_YACScore = X2017_YACScore,
         Y1_YACScore = X2018_YACScore,
         Y0_Overall = X2017_Overall,
         Y1_Overall = X2018_Overall) %>%
  select(-X2017_Team, -X2018_Team)



fulldata_deltas = rbind(wrscore_21_22, wrscore_20_21, wrscore_19_20, wrscore_18_19, wrscore_17_18)
```

Split into train/test
```{r}
sample <- sample(c(TRUE, FALSE), nrow(x_var), replace=TRUE, prob=c(0.8,0.2))
train <- fulldata_deltas[sample, ]
test <- fulldata_deltas[!sample, ]
```


Subset y-variable as Year 1 (vs Year 0) OverallScore
Subset x-variables as all Y0 data and team_change

Linear Regression Model
```{r}
wrscore.lm <- lm(formula = Y1_Overall ~ Y0_Yards + Y0_YardsRun + Y0_Targets + Y0_YardsPerRoute + Y0_CatchScore + Y0_OpenScore + Y0_YACScore + Y0_Overall,
                 data = train)

print(summary(wrscore.lm))

```

Overall it does ok with an R**2 of 0.3441, lets talk out some of the features 

```{r}
wrscore.lm <- lm(formula = Y1_Overall ~ Y0_Yards + Y0_YardsRun + Y0_Targets + Y0_YardsPerRoute + Y0_Overall,
                 data = train)

print(summary(wrscore.lm))

```
Outside of the previous year's overall score, nothing else had a strong prediction for overall 


Lets test it now, assuming we do know the stats from the year. Even if we take away the 3 subset scores, there is probably some correlation among the data.

Ex. Catch Score: ""For example, if the tracking data indicates a pass will be completed 75% of the time and the receiver actually catches the pass, he is credited with plus-0.25. If he does not catch the pass, he is debited at minus-0.75."" So with this, more targets should add to overall score as a whole.


```{r}
catchscore.lm <- lm(formula = Y1_CatchScore ~ Y1_Yards + Y1_YardsRun + Y1_Targets + Y1_YardsPerRoute,
                 data = train)

print(summary(catchscore.lm))


openscore.lm <- lm(formula = Y1_OpenScore ~ Y1_Yards + Y1_YardsRun + Y1_Targets + Y1_YardsPerRoute,
                 data = train)

print(summary(openscore.lm))

yacscore.lm <- lm(formula = Y1_YACScore ~ Y1_Yards + Y1_YardsRun + Y1_Targets + Y1_YardsPerRoute,
                 data = train)

print(summary(yacscore.lm))
```

Specifically 
- Yards and Targets are highly correlated to CatchScore
- Targets and YardsPerRoute is correlated to OpenScore
- Yards and Targets is highly correlated to YACScore



Using all doesn't lead to anything clear, lets do some feature selection

Team Change Variable

```{r}
gg <- ggplot(fulldata_deltas, aes(Y0_Overall, Y1_Overall, colour = as.factor(team_change))) + 
  geom_point() + 
  geom_smooth(alpha=0.3, method="lm") + 
  labs(title = "Year 0 Overall WR Score vs Year 1 Overall WR Score",
       subtitle = "Players who did not change teams had a more consistent 
       catch score on average compared to players who did change.")
print(gg)
```

If there isnt a clear correlation for all the WRs together, lets see if we can create groupings to see if there are patterns among those

-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

We'll use decision trees to analyze it as a whole to start

Decision Tree
```{r}
library(tidymodels)

tree_spec <- decision_tree() %>%
 set_engine("rpart") %>%
 set_mode("regression")

# Fit the model to the training data
tree_fit <- tree_spec %>%
 fit(formula = Y1_Overall ~ Y0_Yards + Y0_YardsRun + Y0_Targets + Y0_YardsPerRoute + Y0_CatchScore + Y0_OpenScore + Y0_YACScore + Y0_Overall,
                 data = train)

# Make predictions on the testing data
predictions <- tree_fit %>%
 predict(test) %>%
 pull(.pred)

# Calculate RMSE and R-squared
metrics <- metric_set(rmse, rsq)
model_performance <- test %>%
 mutate(predictions = predictions) %>%
 metrics(truth = Y1_Overall, estimate = predictions)

print(model_performance)
```
Even though it isnt great (a good amount worse than the linear regression model, lets see how it visualizes)

```{r}
# Load the library
library(rpart.plot)

# Plot the decision tree
rpart.plot(tree_fit$fit, type = 5, extra = 101, under = TRUE, cex = 0.8, box.palette = "auto")
```

Try it without the advanced stats
```{r}
tree_spec <- decision_tree() %>%
 set_engine("rpart") %>%
 set_mode("regression")

# Fit the model to the training data
tree_fit <- tree_spec %>%
 fit(formula = Y1_Overall ~ Y0_Yards + Y0_YardsRun + Y0_Targets + Y0_YardsPerRoute,
                 data = train)

# Make predictions on the testing data
predictions <- tree_fit %>%
 predict(test) %>%
 pull(.pred)

# Calculate RMSE and R-squared
metrics <- metric_set(rmse, rsq)
model_performance <- test %>%
 mutate(predictions = predictions) %>%
 metrics(truth = Y1_Overall, estimate = predictions)

print(model_performance)


# Plot the decision tree
rpart.plot(tree_fit$fit, type = 5, extra = 101, under = TRUE, cex = 0.8, box.palette = "auto")
```

Even though the model treats output as ints / categories rather than continuous, the model generally tells us that yards gained and yards per route are pretty big indicators for a reciever's overall score.




-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------



Now we'll use Regression Trees to create regressions for individual groups 

```{r}
library(rsample)     # data splitting 
library(dplyr)       # data wrangling
library(rpart)       # performing regression trees
library(rpart.plot)  # plotting regression trees
library(ipred)       # bagging
library(caret)       # bagging
```
```{r}
model1 <- rpart(
   formula = Y1_Overall ~ Y0_Yards + Y0_YardsRun + Y0_Targets + Y0_YardsPerRoute + Y0_CatchScore + Y0_OpenScore + Y0_YACScore + Y0_Overall,
   data = train,
   method = 'anova'
)
rpart.plot(model1)

```

```{r}
plotcp(model1)
```

```{r}
model2 <- rpart(
    formula = Y1_Overall ~ Y0_Yards + Y0_YardsRun + Y0_Targets + Y0_YardsPerRoute + Y0_CatchScore + Y0_OpenScore + Y0_YACScore + Y0_Overall,
   data = train,
   method = 'anova',
    control = list(cp = 0, xval = 3)
)

plotcp(model2)
abline(v = 5, lty = "dashed")

model2$cptable

```


