---
title: "CatchScoreEDA"
output: html_document
date: "2023-08-10"
---

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(reshape2) 

```

Data was pulled manually from FiveThirtyEight 
-- pulled into Excel, created a pivot table with all of the players as rows and columns for each stat for each year 
-- Google Sheets Link: https://docs.google.com/spreadsheets/d/1PI-LMohmf9-yC4w9Fq4XyVQqt_umJd9Bau_nqpVD5Js/edit?usp=sharing


```{r}
data = read.csv("../Data/WR_Advanced_Stats - PivotData_Clean.csv")
head(data)
```

## Which WRs had the biggest change in Catch Score from 2021 to 2022? 
```{r}
wrscore_21_22 = data %>%
  select(Player, X2021_Overall, X2022_Overall) %>%
  mutate(overall_delta = X2022_Overall - X2021_Overall) %>%
  mutate(overall_delta_abs = abs(overall_delta)) %>%
  na.exclude() %>%
  arrange(desc(overall_delta_abs))
  
  
wrscore_21_22 %>%
  head(15) %>%
  mutate(color = ifelse(overall_delta < 0, "#FFDDDD","#CCEEFF")) %>%
  ggplot(aes(x = reorder(Player, -overall_delta), y = overall_delta, fill = color, label = overall_delta)) + 
    geom_bar(stat="identity") +
    geom_label() +
    theme_minimal() +
    labs(title = "Top 15 Biggest Delta in Overall WR Score from 2021 to 2022",
         x = "Player", 
         y = "Change in WR Score 2021-2022") + 
    theme(axis.text.x = element_text(angle = 30, vjust = 0.7)) +
    scale_fill_identity(guide = FALSE)

ggsave("../Images/BiggestCatchScoreDelta_2021v2022.png", bg="white")

```

How do all WRs look 2021 to 2022?
```{r}
library(ggrepel)


wrscore_21_22 %>%
  ggplot(aes(x = X2021_Overall, y = X2022_Overall, label = Player)) + 
    geom_point(color = 'lightblue' ) +
    geom_text_repel(aes(label=ifelse(overall_delta_abs>16,as.character(Player),'')), vjust=0,hjust="inward",size = 3) +
    geom_abline(slope=1, intercept = 0, color = 'lightgreen') +
    geom_text(aes(30, 12, label = "Line for 2021 Score = 2022 Score (Y=x line)"), color = 'darkgreen') +
    theme_minimal() + 
    labs(title = "Biggest Delta in Overall WR Score from 2021 to 2022",
         subtitle = "Players labeled have Overall Catch Score Delta greater than 16",
         x = "2021 Catch Score", 
         y = "2022 Catch Score")

ggsave("../Images/CatchScore_2021v2022.png", bg="white")

```

Overall, players who have a high score keep consistent catch scores, unless a player deals with injury (Andrews, Kupp) or situations changed and they became more of a primary weapon on their team, limiting their catch ability (Adams, Kirk)


## Which WRs had the biggest change in Catch Score from any two years?


```{r}
data_overalls = data %>%
  select(Player, X2017_Overall, X2018_Overall, X2019_Overall, X2020_Overall, X2021_Overall, X2022_Overall)

data_overalls$max_overall <- apply(data_overalls[, 2:7], 1, max, na.rm=TRUE)
data_overalls$min_overall <- apply(data_overalls[, 2:7], 1, min, na.rm=TRUE)


wrscore_alldelta = data_overalls %>%
  mutate(overall_delta = max_overall - min_overall) %>%
  mutate(overall_delta_abs = abs(overall_delta)) %>%
  arrange(desc(overall_delta_abs))
  
  
wrscore_alldelta %>%
  head(15) %>%
  mutate(color = ifelse(overall_delta < 0, "#FFDDDD","#CCEEFF")) %>%
  ggplot(aes(x = reorder(Player, -overall_delta), y = overall_delta, fill = '#CBC3E3', label = overall_delta)) + 
    geom_bar(stat="identity") +
    geom_label() +
    theme_minimal() +
    labs(title = "Top 15 Biggest Delta in Overall WR Score between any two years (17-22)",
         x = "Player", 
         y = "Diff. of Max WR Score and Min WR Score") + 
    theme(axis.text.x = element_text(angle = 30, vjust = 0.7)) +
    scale_fill_identity(guide = FALSE)

ggsave("../Images/BiggestCatchScoreDelta_2017_2022.png", bg = 'white')

```

Key Insights
- A lot of top WRs here who either had an amazing season, and then had a drastic fall in WR score in following years (Antonio Brown, Julio Jones, Michael Thomas); or had a great season compared to their rookie years (Amari Cooper, Zay Jones)
- A lot of TEs here, where opportunity and involvement in the pass game was more of a clear factor


## Which WRs had the biggest change in Catch Score from two CONSECUTIVE years?

```{r}
wrscore_21_22 = wrscore_21_22 %>%
  rename(Y0_Overall = X2021_Overall,
         Y1_Overall = X2022_Overall) %>%
  mutate(Player = paste(Player, "21-22"))

wrscore_20_21 = data %>%
  select(Player, X2020_Overall, X2021_Overall) %>%
  mutate(overall_delta = X2021_Overall - X2020_Overall) %>%
  mutate(overall_delta_abs = abs(overall_delta)) %>%
  na.exclude() %>%
  rename(Y0_Overall = X2020_Overall,
         Y1_Overall = X2021_Overall) %>%
  mutate(Player = paste(Player, "20-21"))

wrscore_19_20 = data %>%
  select(Player, X2019_Overall, X2020_Overall) %>%
  mutate(overall_delta = X2020_Overall - X2019_Overall) %>%
  mutate(overall_delta_abs = abs(overall_delta)) %>%
  na.exclude() %>%
  rename(Y0_Overall = X2019_Overall,
         Y1_Overall = X2020_Overall) %>%
  mutate(Player = paste(Player, "19-20"))

wrscore_18_19 = data %>%
  select(Player, X2018_Overall, X2019_Overall) %>%
  mutate(overall_delta = X2019_Overall - X2018_Overall) %>%
  mutate(overall_delta_abs = abs(overall_delta)) %>%
  na.exclude() %>%
  rename(Y0_Overall = X2018_Overall,
         Y1_Overall = X2019_Overall) %>%
  mutate(Player = paste(Player, "18-19"))

wrscore_17_18 = data %>%
  select(Player, X2017_Overall, X2018_Overall) %>%
  mutate(overall_delta = X2018_Overall - X2017_Overall) %>%
  mutate(overall_delta_abs = abs(overall_delta)) %>%
  na.exclude() %>%
  rename(Y0_Overall = X2018_Overall,
         Y1_Overall = X2017_Overall) %>%
  mutate(Player = paste(Player, "17-18"))


wrscore_deltas = rbind(wrscore_21_22, wrscore_20_21, wrscore_19_20, wrscore_18_19, wrscore_17_18) %>%
    arrange(desc(overall_delta_abs))
```

```{r}
wrscore_deltas %>%
  head(20) %>%
  mutate(color = ifelse(overall_delta < 0, "#FFDDDD","#CCEEFF")) %>%
  ggplot(aes(x = reorder(Player, -overall_delta), y = overall_delta, fill = color, label = overall_delta)) + 
    geom_bar(stat="identity") +
    geom_label() +
    theme_minimal() +
    labs(title = "Top 15 Biggest Delta in Overall WR Scores",
         caption = "Exp: First bar is Amari Cooper's 2018 WR Score minus 2017 WR Score",
         x = "Player (Two Seasons for Delta)", 
         y = "Change in WR Scores") + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.7)) +
    scale_fill_identity(guide = FALSE)

ggsave("../Images/BiggestCatchScoreDelta_consecutive.png", bg = 'white')

```






-------------------------------------------------------------------------------------------------------------------------------------------------
Filter for only players with at least two seasons
```{r}
data_overall = data %>%
  select(Player, X2017_Overall, X2018_Overall, X2019_Overall, X2020_Overall, X2021_Overall, X2022_Overall)

#get consecutive years 
data_overall$consecutive_years <- (apply(!is.na(data_overall[-1]),1,
            function(x) any(with(rle(x), lengths[values])>1)))+0

consecutiveyears_players = data_overall %>%
  filter(consecutive_years > 0) %>%
  select(Player) 

nrow(consecutiveyears_players)
```

The number of players that have at least two consecutive seasons with the threshold of targets is 141



